---
title: "EU LFS Data Management"
author: "Zbignev Gricevic"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Uploading packages

Uploading pre-installed packages needed for data management of EU SILC public use files.

```{r echo=TRUE, warning=FALSE, message=FALSE}
library(data.table)
```

# Working Directory 

Setting working directory.

```{r echo=TRUE, warning=FALSE}
setwd("/home/gricezb/data/LFS/")
```

# Creating Cumulative LFS Country Files

List of countries to be read in.

```{r echo=TRUE, warning=FALSE}
country<-c("AT", "BE", "BG", "CY", "DE", "DK", "EE", "EL", "ES", "FI", "FR", "HU", "IE", "IT", "LT", "LU", "LV", "MT", "NL", "PT", "RO", "SI", "SK")
year<-c(2004:2013)
```

# Creating Cumulative LFS Country files

Creating list of files to be merged. First of all list for each country is
created. Then these lists are combined into one large list (lfs_files).

```{r echo=TRUE, warning=FALSE}
for (i in country){
  assign(paste0("temp_", i), list.files(pattern = paste0(i, "_LFS_", ".*")))
}
lfs_files<-mget(ls(pattern = "temp.*"))
```

Loop is defined which creates and saves cumulative LFS file for each country.
We first read files for a specific country. Then we bind all country files into
one data table (country code_files_merged). In the next step we export 
country files to the .rds and .csv formats.

```{r echo=TRUE, warning=FALSE}
for (i in lfs_files){
  assign(substr(i, 1, 2), lapply(i, fread))
  assign(paste0(substr(i, 1, 2), "_files_merged"), rbindlist(get(paste0(substr(i, 1, 2))),  fill = TRUE))
  rm(list = substr(i, 1, 2))
  gc()
  saveRDS(get(paste0(substr(i, 1, 2), "_files_merged")), paste0(substr(i[1], 1, 2), "_files_merged.rds"))
  rm(list=ls(pattern="_files_merged"))
  gc()
  }
```

# Creating Cumulative LFS Yearly files

Creating list of files to be merged. First of all list for each year is
created. Then these lists are combined into one large list (lfs_files).

```{r echo=TRUE, warning=FALSE}
for (i in year){
  assign(paste0("temp_", i), list.files(pattern = paste0("*", "_LFS_", i , ".*")))
}
lfs_files<-mget(ls(pattern = "temp.*"))
```

Loop is defined which creates and saves cumulative LFS file for each year.
We first read files for a specific year. Then we bind all year files into
one data table (year code_files_merged). In the next step we export 
yearly files to the .rds and .csv format.

```{r echo=TRUE, warning=FALSE}
for (i in lfs_files){
  assign(substr(i, 8, 11), lapply(i, fread))
  assign(paste0(substr(i, 8, 11), "_files_merged"), rbindlist(get(paste0(substr(i, 8, 11))),  fill = TRUE))
  rm(list = substr(i, 1, 2))
  gc()
  saveRDS(get(paste0(substr(i, 8, 11), "_files_merged")), paste0(substr(i[1], 8, 11), "_files_merged.rds"))
  rm(list=ls(pattern="_files_merged"))
  gc()
}
```

# Creating Cumulative LFS file

```{r echo=TRUE, warning=FALSE}
lfs_files <- list.files(pattern = "merged")
```
Define the loop which uploads and appends country files
iteratively. After programme appends country file to the 
cumulative file, it removes it from the Enviroment.

```{r echo=TRUE, warning=FALSE}
for (file in lfs_files){
if (!exists("all_lfs")){
    all_lfs <- NULL
  }
if (exists("all_lfs")){
    temp_dataset <-readRDS(file)
    all_lfs<-rbindlist(list(all_lfs, temp_dataset), fill=TRUE)
    rm(temp_dataset)
    gc()
  } 
}
```
Cross-check observations per country.

```{r echo=TRUE, warning=FALSE}
table(all_lfs$COUNTRY)
```
