---
title: "EU LFS Data Management"
author: "Zbignev Gricevic"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Uploading packages

Uploading pre-installed packages needed for data management of EU SILC public use files.

```{r echo=TRUE, warning=FALSE, message=FALSE}
packages<-c("data.table")
lapply(packages, library, character.only = TRUE)
```

# Working Directory 

Setting working directory.

```{r echo=TRUE, warning=FALSE}
setwd("/home/gricezb/data/LFS/")
```

# Creating Cumulative LFS Country Files

List of countries to be read in.

```{r echo=TRUE, warning=FALSE}
#country<-c("AT", "BE", "BG", "CY", "DE", "DK", "EE", "EL", "ES", "FI", "FR", "HU", "IE", "IT", "LT", "LU", "LV", "MT", "NL", "PT", "RO", "SI", "SK")
country<-c("AT", "BE")

```

Creating list of files to be merged. First of all list for each country is
created. Then these lists are combined into one large list (lfs_files).

```{r echo=TRUE, warning=FALSE}
for (i in country){
  assign(paste0("temp_", i), list.files(pattern = paste0(i, "_LFS_", ".*")))
}
lfs_files<-mget(ls(pattern = "temp.*"))
```

Loop is defined which creates and saves cumulative LFS file for each country.
We first read files for a specific country. Then we bind all country files into
one data table ('country code'_files_merged). In the next step we export 
country files to the .rds format.

```{r echo=TRUE, warning=FALSE}
for (i in lfs_files){
  assign(substr(i, 1, 2), lapply(i, fread))
  assign(paste0(substr(i, 1, 2), "_files_merged"), rbindlist(get(paste0(substr(i, 1, 2))),  fill = TRUE))
  rm(list = substr(i, 1, 2))
  gc()
  saveRDS(get(paste0(substr(i, 1, 2), "_files_merged")), paste0(substr(i[1], 1, 2), "_files_merged.rds"))
  rm(list=ls(pattern="_files_merged"))
  gc()
  }
```

# Creating Cumulative LFS File

```{r echo=TRUE, warning=FALSE}
lfs_files <- list.files(pattern = "merged")
```

Define the loop which uploads and appends country files
iteratively. After programme appends country file to the 
cumulative file, it removes it from the Enviroment.

```{r echo=TRUE, warning=FALSE}
for (file in lfs_files){
if (!exists("all_lfs")){
    all_lfs <- NULL
  }
if (exists("all_lfs")){
    temp_dataset <-readRDS(file)
    all_lfs<-rbindlist(list(all_lfs, temp_dataset), fill=TRUE)
    rm(temp_dataset)
    gc()
  } 
}
```


```{r echo=TRUE, warning=FALSE}
table(all_lfs$COUNTRY)
```
