---
title: "EU SILC PUF Data Management"
author: "Zbignev Gricevic"
date: "19 février 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Uploading packages

Uploading pre-installed packages needed for data management of EU SILC public use files.

```{r echo=TRUE, warning=FALSE, message=FALSE}
packages<-c("plyr", "dplyr", "colr", "ggplot2", "plotly", "data.table")
lapply(packages, library, character.only = TRUE)
```

# Working Directory 

Setting working directory.

```{r echo=TRUE, warning=FALSE}
setwd("/home/gricezb/data/EU Silc/")
```

#Household Files

Uploading and merging household register and household files.

##Household Register (d) files.

Creating list of household register files (temp_d) stored in the 
working directory. Reading in the files into working Environment. 
Binding files into one large list (d_files). Removing objects used
to create this list from the working environment.

```{r echo=TRUE, warning=FALSE, message=FALSE}
temp_d=list.files(pattern = "*d_EUSILC.csv")
for (i in 1:length(temp_d)) assign(temp_d[i], fread(temp_d[i]))
d_files<-mget(ls(pattern = "*d_EUSILC.csv"))
rm(list=ls(pattern="d_EUSILC.csv"), temp_d)
invisible(gc())
```

Appending list of household register files (d_files) into single 
data table (d_files_all). Removing list of files from the working
environment.

```{r echo=TRUE, warning=FALSE}
d_files_all<-rbindlist(d_files,  fill = TRUE)
rm(d_files)
invisible(gc())
```

One country (Greece) has two different names in the variable storing 
country codes (DB020). For some years Greece is coded as "GR" and for 
other years as "EL". To reduce confusion we recode all instances of "GR"
to "EL".

```{r echo=TRUE, warning=FALSE}
d_files_all$DB020[d_files_all$DB020=="GR"] <- "EL"
```

Exporting household register file to working directory and temporarily
removing it from working environment.

```{r echo=TRUE, warning=FALSE}
saveRDS(d_files_all, "d_files_all.rds")
rm(d_files_all)
invisible(gc())
```

##Household (h) files.

Creating list of household files (temp_h) stored in the 
working directory. Reading in the files into working Environment. 
Binding files into one large list (h_files). Removing objects used
to create this list from the working environment.

```{r echo=TRUE, warning=FALSE, message=FALSE}
temp_h=list.files(pattern = "*h_EUSILC.csv")
for (i in 1:length(temp_h)) assign(temp_h[i], fread(temp_h[i]))
h_files<-mget(ls(pattern = "*h_EUSILC.csv"))
rm(list=ls(pattern="h_EUSILC.csv"), temp_h)
invisible(gc())
```

Appending list of household files (h_files) into single 
data table (h_files_all). Removing list of files from the working
environment.

```{r echo=TRUE, warning=FALSE}
h_files_all<-rbindlist(h_files,  fill = TRUE)
rm(h_files)
invisible(gc())
```

One country (Greece) has two different names in the variable storing 
country codes (HB020). For some years Greece is coded as "GR" and for 
other years as "EL". To reduce confusion we recode all instances of "GR"
to "EL".

```{r echo=TRUE, warning=FALSE}
h_files_all$HB020[h_files_all$HB020=="GR"] <- "EL"
```

Exporting household file to working directory and temporarily
removing it from working environment.

```{r echo=TRUE, warning=FALSE}
saveRDS(h_files_all, "h_files_all.rds")
rm(h_files_all)
invisible(gc())
```

##Merging Household Register and Household Files

Uploading household register (d_files_all) and household (h_files_all) cumulative files.

```{r echo=TRUE, warning=FALSE}
d_files_all<-readRDS("d_files_all.rds")
h_files_all<-readRDS("h_files_all.rds")
```

Information on Year, Country and Household ID will be used to merge household register (d_files_all) 
and household (h_files_all) cumulative files. Renaming these variables to have common names in both
data.tables.

```{r echo=TRUE, warning=FALSE}
d_files_all<-setnames(d_files_all, old=c("DB010","DB020", "DB030"), new=c("Year", "Country", "HHID"))
h_files_all<-setnames(h_files_all, old=c("HB010","HB020", "HB030"), new=c("Year", "Country", "HHID"))
```

Merging household register and household files into cumulative file. Removing files used to
do merging from the working environment.

```{r echo=TRUE, warning=FALSE}
household_file <- merge(d_files_all, h_files_all, by = c("Country", "HHID", "Year"), all = TRUE)
rm(d_files_all, h_files_all)
invisible(gc())
```

Exporting cumulative household file to working directory and temporarily
removing it from working environment.

```{r echo=TRUE, warning=FALSE}
saveRDS(household_file, "household_file.rds")
rm(household_file)
invisible(gc())
```

#Personal Files

Uploading and merging personal register and personal files.

##Personal Register (r) files.

Creating list of personal register files (temp_r) stored in the 
working directory. Reading in the files into working Environment. 
Binding files into one large list (r_files). Removing objects used
to create this list from the working environment.

```{r echo=TRUE, warning=FALSE, message=FALSE}
temp_r=list.files(pattern = "*r_EUSILC.csv")
for (i in 1:length(temp_r)) assign(temp_r[i], fread(temp_r[i]))
r_files<-mget(ls(pattern = "*r_EUSILC.csv"))
rm(list=ls(pattern="r_EUSILC.csv"), temp_r)
invisible(gc())
```

Appending list of household register files (r_files) into single 
data table (r_files_all). Removing list of files from the working
environment.

```{r echo=TRUE, warning=FALSE}
r_files_all<-rbindlist(r_files,  fill = TRUE)
rm(r_files)
invisible(gc())
```

One country (Greece) has two different names in the variable storing 
country codes (RB020). For some years Greece is coded as "GR" and for 
other years as "EL". To reduce confusion we recode all instances of "GR"
to "EL".

```{r echo=TRUE, warning=FALSE}
r_files_all$RB020[r_files_all$RB020=="GR"] <- "EL"
```

A list of variables is named in uppercase in some years and in lowercase
in other years. To reduce confusion, these variables are renamed to uppercase.

```{r echo=TRUE, warning=FALSE}
colnames(r_files_all)[colnames(r_files_all)=="rl010"] <- "RL010"
colnames(r_files_all)[colnames(r_files_all)=="rl020"] <- "RL020"
colnames(r_files_all)[colnames(r_files_all)=="rl030"] <- "RL030"
colnames(r_files_all)[colnames(r_files_all)=="rl040"] <- "RL040"
colnames(r_files_all)[colnames(r_files_all)=="rl050"] <- "RL050"
colnames(r_files_all)[colnames(r_files_all)=="rl060"] <- "RL060"
colnames(r_files_all)[colnames(r_files_all)=="rl070"] <- "RL070"
colnames(r_files_all)[colnames(r_files_all)=="rl010_f"] <- "RL010_F"
colnames(r_files_all)[colnames(r_files_all)=="rl020_f"] <- "RL020_F"
colnames(r_files_all)[colnames(r_files_all)=="rl030_f"] <- "RL030_F"
colnames(r_files_all)[colnames(r_files_all)=="rl040_f"] <- "RL040_F"
colnames(r_files_all)[colnames(r_files_all)=="rl050_f"] <- "RL050_F"
colnames(r_files_all)[colnames(r_files_all)=="rl060_f"] <- "RL060_F"
colnames(r_files_all)[colnames(r_files_all)=="rl070_f"] <- "RL070_F"
```

Exporting household register file to working directory and temporarily
removing it from working environment.

```{r echo=TRUE, warning=FALSE}
saveRDS(r_files_all, "r_files_all.rds")
rm(r_files_all)
invisible(gc())
```

##Personal (p) files.

Creating list of personal register files (temp_p) stored in the 
working directory. Reading in the files into working Environment. 
Binding files into one large list (p_files). Removing objects used
to create this list from the working environment.

```{r echo=TRUE, warning=FALSE, message=FALSE}
temp_p=list.files(pattern = "*p_EUSILC.csv")
for (i in 1:length(temp_p)) assign(temp_p[i], fread(temp_p[i]))
p_files<-mget(ls(pattern = "*p_EUSILC.csv"))
rm(list=ls(pattern="p_EUSILC.csv"), temp_p)
invisible(gc())
```

Appending list of personal files (p_files) into single 
data table (p_files_all). Removing list of files from the working
environment.

```{r echo=TRUE, warning=FALSE}
p_files_all<-rbindlist(p_files,  fill = TRUE)
rm(p_files)
invisible(gc())
```

One country (Greece) has two different names in the variable storing 
country codes (PB020). For some years Greece is coded as "GR" and for 
other years as "EL". To reduce confusion we recode all instances of "GR"
to "EL".

```{r echo=TRUE, warning=FALSE}
p_files_all$PB020[p_files_all$PB020=="GR"] <- "EL"
```

A list of variables is named in uppercase in some years and in lowercase
in other years. To reduce confusion, these variables are renamed to uppercase.

```{r echo=TRUE, warning=FALSE}
colnames(p_files_all)[colnames(p_files_all)=="pl030"] <- "PL030"
colnames(p_files_all)[colnames(p_files_all)=="pl070"] <- "PL070"
colnames(p_files_all)[colnames(p_files_all)=="pl072"] <- "PL072"
colnames(p_files_all)[colnames(p_files_all)=="pl030_f"] <- "PL030_F"
colnames(p_files_all)[colnames(p_files_all)=="pl070_f"] <- "PL070_F"
colnames(p_files_all)[colnames(p_files_all)=="pl072_f"] <- "PL072_F"
colnames(p_files_all)[colnames(p_files_all)=="pl110_f"] <- "PL110_F"
```

Exporting personal file to working directory and temporarily
removing it from working environment.

```{r echo=TRUE, warning=FALSE}
saveRDS(p_files_all, "p_files_all.rds")
rm(p_files_all)
invisible(gc())
```

##Merging Personal Register and Personal Files

Uploading personal register (r_files_all) and personal (p_files_all) cumulative files.

```{r echo=TRUE, warning=FALSE}
r_files_all<-readRDS("r_files_all.rds")
p_files_all<-readRDS("p_files_all.rds")
```

Information on Year, Country and Personal ID will be used to merge personal register (r_files_all) 
and personal (p_files_all) cumulative files. Renaming these variables to have common names in both
data.tables.

```{r echo=TRUE, warning=FALSE}
r_files_all<-setnames(r_files_all, old=c("RB010","RB020", "RB030", "RX030"), new=c("Year", "Country", "PID", "HHID"))
p_files_all<-setnames(p_files_all, old=c("PB010","PB020", "PB030", "PX030"), new=c("Year", "Country", "PID", "HHID"))
```

Merging personal register and personal files into cumulative file. Removing files used to
do merging from the working environment.

```{r echo=TRUE, warning=FALSE}
personal_file <- merge(r_files_all, p_files_all, by = c("Country", "PID", "Year"), all = TRUE)
rm(r_files_all, p_files_all)
invisible(gc())
```

Exporting cumulative personal file to working directory and temporarily
removing it from working environment.

```{r echo=TRUE, warning=FALSE}
saveRDS(personal_file, "personal_file.rds")
rm(personal_file)
invisible(gc())
```

#Cumulative EU SILC file

Uploading household and personal cumulative files.

```{r echo=TRUE, warning=FALSE}
household_file<-readRDS("household_file.rds")
personal_file<-readRDS("personal_file.rds")
```

Merging household and personal cumulative files into cumulative file. Removing files used to
do merging from the working environment.

```{r echo=TRUE, warning=FALSE}
EU_SILC_Cumulative <- merge(household_file, personal_file, by = c("Country", "HHID", "PID", "Year"), all = TRUE)
```

```{r echo=TRUE, warning=FALSE}
saveRDS(EU_SILC_Cumulative, "EU_SILC_Cumulative.rds")
rm(EU_SILC_Cumulative)
invisible(gc())
```

